Predicting instabilities of a tuneable ring laser with an iterative map model

### 2020-11-22 Papers/Journal/Tuneable_Lasers.tex
Fixed energy and power section

\section{Introduction}
\label{sec:intro}
Tuneable lasers have the ability to vary the frequency of their output by up to about 100 nanometres \cite{bohun2015, burgoyne2010, yamashita2009}. This ability to vary the frequency leads to applications such as optical coherence tomography \cite{bohun2015, burgoyne2014, yamashita2009}, coherent anti-Stokes Raman spectroscopy \cite{burgoyne2014}, deep tissue multi-photon microscopy \cite{chung2017, liu2017}, and diagnostics of ultrafast processes \cite{burgoyne2014, silfvast2004}. Tuneable lasers are usually constructed in a ring with five main components: output coupler, chirped fibre Bragg grating (CFBG), modulator, Er-doped fibre, and pump laser. A typical ring laser cavity is depicted in Fig. \ref{fig:cavity}. Due to the high power and short duration of the pulses, the Kerr nonlinearity plays a crucial role in the dynamics within the cavity.

Several effects arise within the cavity due to the interplay of dispersion, modulation, and the nonlinearity \cite{bohun2015, coen1997, lapre2019, shao2019, woodward2018}. The two effects of most interest for this paper are wave breaking, and modulation instability. Wave breaking causes the leading edge of a pulse to be red shifted, and the trailing edge blue shifted (in the anomalous dispersion regime, the effect is reversed) to the point that a shock is developed \cite{anderson1992, rothenberg1989a, rothenberg1989b, tomlinson1984, tomlinson1985}. The shifts in frequencies cause the pulse to become more rectangular in the frequency domain with a linear chirp over most of the pulse. In optics, wave breaking manifests from self-phase modulation (SPM) which is a direct effect of the Kerr nonlinearity causing the pulse to interfere with itself. New frequencies are then generated which induces the rectangular profile \cite{agrawal2013, mollenauer1980, woodward2018}. Outside of optics, wave breaking occurs in other areas with nonlinear waves such as plasmas, heat propagation, transmission lines, and fluid dynamics \cite{rothenberg1989b}. Modulation instability---which typically arises in the anomalous dispersion regime---is the other effect that will be of interest. However, in the presence of multiple frequencies, modulation instability can emerge in the normal dispersion regime as well through cross-phase modulation and four-wave mixing \cite{agrawal1987, agrawal1989, agrawal2013, haelterman1992}. Modulation instability causes a wave to break into short pulses, and can cause a laser to lose coherence \cite{agrawal1987, coen1997, haelterman1992}. Within a ring laser cavity wave breaking and modulation instability can become parasitic leading to an unstable and unsustainable pulse. This gives rise to the need of understanding the rich landscape of the parameter space, and determining design principles to guarantee the ring laser is stable and sustainable \cite{bohun2015, burgoyneemail, finot2008, lapre2019, woodward2018}.

\begin{figure}[tbp]
	\centering
	\input{Cavity}
	\caption{Typical cavity of a fibre ring laser \cite{burgoyne2014, chung2017, lapre2019, shao2019, tang2014}. The laser pulses travel clockwise around each loop.}
	\label{fig:cavity}
\end{figure}

In this paper, we will build an iterative map model for the evolution of a laser pulse as it travels though the cavity. In Section \ref{sec:modelling}, we will briefly review previous mathematical models for ring lasers. Then, we will present our own iterative map model in Section \ref{sec:model}, as well as non-dimensionalize the model and introduce four dimensionless parameters that govern the dynamics of the system. We will then examine the results of our model in Section \ref{sec:results}---first solving the model analytically, by assuming the absence of the nonlinearity, then numerically solving the full nonlinear model. We will identify a sharp boundary in the parameter space separating the region of stability, and the region where modulation instability degrades the pulse. Finally, the concluding thoughts and possible ideas for future work are given in Section \ref{sec:conclusion}.

\section{Modelling Efforts}
\label{sec:modelling}
The standard equation for studying nonlinear optics is the nonlinear Schr\"odinger equation (NLSE),
\begin{align}
	\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + i \gamma |A|^2 A.
	\label{eq:smallnlse}
\end{align}
Here, $A = A(T, z) : \mathbb{R}^2 \mapsto \mathbb{C}$ is the complex pulse amplitude, $\beta_2 \in \mathbb{R}$ is the second order dispersion, and $\gamma \in \mathbb{R}$ is the coefficient of nonlinearity. In practice, the NLSE, \eqref{eq:smallnlse}, lacks a few key terms, thus, it is often generalized by adding amplification and loss (occasionally higher order terms are added as well). These additions give the generalized nonlinear Schr\"{o}dinger equation (GNLSE) \cite{agrawal2013, bohun2015, finot2008, peng2018, shtyrina2017, yarutkina2013},
	\begin{align}
	\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + i \gamma |A|^2 A + \frac{1}{2}g(A) A - \alpha A,
	\label{eq:nlse}
\end{align}
where $g(A)$ is an amplifying term due to the gain, and $\alpha \in \mathbb{R}^+$ is the loss due to scattering and absorption.

The GNLSE has many applications in nonlinear optics and fibre optic communications, however, in laser physics typically a modulation term is also added to ensure mode-locking, this yields the master equation of mode-locking \cite{haus1984, haus1975, haus1986, haus1992, haus2000, tamura1996, usechak2005},
\begin{align}
	\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + i \gamma |A|^2 A + \frac{1}{2}g(A) A - \alpha A - M(T).
	\label{eq:meml}
\end{align}
No analytic solution is known for \eqref{eq:meml}---even after making the assumptions that the gain is constant, and the modulation is quadratic \cite{haus1984, haus1975, haus1996}. However, by neglecting certain terms of \eqref{eq:meml} analytic solutions can be found \cite{burgoyne2014, haus1975, haus1986, haus1991, haus1992, haus1996, tamura1996, usechak2005}. See Haus \cite{haus2000} for a comprehensive description and history of the theory of mode-locked lasers.

\subsection{Discrete Models}
\label{sec:discrete}
Unfortunately, the master equation, \eqref{eq:meml}, is not completely representative of the underlying physics within our laser cavity. It is assumed in \eqref{eq:meml} that each process affects the pulse continuously within the cavity. As highlighted by Fig. \ref{fig:cavity}, this can be a poor assumption. Within the cavity, each effect is localized to its corresponding component: almost all of the dispersion happens within the CFBG \cite{agrawal2002}, the pulse is only amplified within the Erbium-doped fibre, \emph{et cetera}. Thus, a better model is one where we consider a simplified version of \eqref{eq:meml} for each component. Solving these simplified differential equations yields an algebraic expression for the effect of each individual component. The relations can then be functionally composed to give an iterative map for the effect of one round trip of the cavity.

Such a method was first proposed in 1955 by Cutler \cite{cutler1955} while analyzing a microwave regenerative pulse generator. His method was adapted for mode-locked lasers in 1969 by Siegman and Kuizenga \cite{kuizenga1970a, kuizenga1970b, kuizenga1970, siegman1969}. The effects of the nonlinearity would not be considered until Martinez \emph{et al.} \cite{martinez1984, martinez1985} modelled passively mode-locked lasers. This issue has recently been readdressed by Burgoyne \cite{burgoyne2014} in the literature for tuneable lasers. In each of these models the effect of each block is described by a transfer function.

Despite the development of discrete, block style models, several short-comings exist. None of the previous models have contained every component within a laser cavity---either the nonlinearity or the modulation have been omitted. Each component of a tuneable plays a pivotal role and the laser will not function correctly without the inclusion of all of the components. Another drawback is that the functional operations of some of the components used in their models are phenomenological. While these functions are chosen based on the observed output, they are not necessarily consistent with the underlying physics. Finally, the previous models have been unsuccessful in exhibiting the instabilities described in Section \ref{sec:intro}.

\section{Model Derivation}
\label{sec:model}
Using the idea of a discrete model presented Section \ref{sec:discrete}, we derive our model from the GNLSE, \eqref{eq:nlse}; however, in the case of the modulation we consider the exact functional form to be determined by the laser operator. To accomplish this we follow the method described in Section \ref{sec:discrete}; we neglect all terms but one on the right side of \eqref{eq:nlse} and solve the simplified differential equations.

To proceed we must choose the form of the amplification term, $g(A)$. We assume the gain has the form
\begin{align}
	g(A) &= \frac{g_0}{1 + E / \Es},& E &= \int_{-\infty}^\infty |A|^2 \, \df T,
	\label{eq:energy}
\end{align}
where $g_0$ is a small signal gain, $E$ is the energy of the pulse, and $\Es$ is the energy at which the gain begins to saturate \cite{haus1984, shtyrina2017, silfvast2004}. This reduces \eqref{eq:nlse} to
\begin{align}
	\pdiff{A}{z} &= \frac{g_0 A}{2 \left( 1 + E / \Es \right)}.
\end{align}
By assuming the incoming energy is $E$ and the energy after travelling through the gain fibre of length $L_g$ is $E_{\text{out}}$, we find the energy after amplification is
\begin{align}
	\label{eq:engain}
	E_{\text{out}} &= \Es W \left( \frac{E}{\Es} \exp( E / \Es ) \exp( g_0 L_g ) \right),
\end{align}
where $W$ is the Lambert $W$ function. Rewriting \eqref{eq:engain} in terms of the pulse gives
\begin{align}
	G(A) &= \left( \frac{\Es}{E} W \left( \frac{E}{\Es} \exp( E / \Es ) \exp( g_0 L_g ) \right) \right)^{1/2} A
	\label{eq:gainform}
\end{align}
for the amplification of the pulse within the Er-doped fibre. Repeating this process for the loss, dispersion, and nonlinearity yields
\begin{align}
	L(A) &= (1 - R) \exp( - \alpha L_T )A, \\
	D(A) &= \FTi{\exp( i \omega^2 L_D\beta_2/2 ) \FT{A}}, \\
	F(A) &= A \exp( i \gamma |A|^2 L_f ),
\end{align}
where $R$ is the reflectivity of the output coupler (depending on the layout of the laser cavity the loss may instead take the form $L(A) = R \exp( - \alpha L_T )A$), $L_T$ is the total length of the laser circuit, $L_D$ is the length of the dispersive medium, $L_f$ is the length of fibre between the amplifier and the output coupler, and $\mathcal{F}$ denotes the Fourier transform. Finally, we must consider the modulation. The modulation is considered to be applied externally in which ever way the operator sees fit. For simplicity the representation is taken as the Gaussian
\begin{align}
	M(A) &= \exp( -T^2 / 2 T_M^2 ) A,
	\label{eq:modform}
\end{align}
where $T_M$ is the characteristic width of the modulation. This assumption is not particularly restrictive. Calcaterra and Boldt have shown that Gaussians can form a basis for $L^2(\mathbb{R})$ \cite{calcaterra2008a}. Therefore, we can approximate any modulation function in $L^2(\mathbb{R})$ by a sum of \eqref{eq:modform}.

\subsection{Non-Dimensionalization}
Non-dimensionalization is a technique used in mathematical modelling to better understand the relative importance of different processes within a system. In essence, we scale all variables by typical values of the system based on the geometry and boundary conditions (see, for example, \cite{howison2005}). Doing so `factors' the dimensional units out of the problem, leaving only dimensionless equations with dimensionless parameters. The magnitude of the dimensionless parameters characterizes the dominant dynamics of the system---much like how fluid flow can be described with only the Reynolds number. The structure of each process in the laser can be better understood by re-scaling the time, energy, and amplitude by the convenient scalings:
\begin{align}
	T &= T_M \widetilde{T},& E &= \Es \widetilde{E},& A &= \left( \frac{\Es}{T_M} \right)^{1/2} \widetilde{A}.
\end{align}

Revisiting the process maps \eqref{eq:gainform}--\eqref{eq:modform} yields the new, non-dimensional, mappings---after dropping the tildes---
\begin{align}
	G(A) &= \left(E^{-1} W \left( a E \textrm{e}^{E}\right) \right)^{1/2} A, & F(A) &= A \exp( i b |A|^2), \nonumber \\
	D(A) &= \FTi{\exp( i s^2 \omega^2 ) \FT{A}}, & L(A) &= h A, \label{eq:effects} \\
	M(A) &= \exp( -T^2 / 2 ) A, \nonumber
\end{align}
with the four dimensionless parameters, as defined by the values in Table \ref{tab:values},
\begin{equation}
	\begin{aligned}
		a &= \exp( g_0 L_g ) \sim 8 \times 10^3, & 
		h &= (1 - R) \exp( - \alpha L_T) \sim 0.04, \\
		b &= \gamma L_f \frac{\Es}{T_M} \sim 1, & s &= \left( \frac{\beta_2 L_D}{2 T_M^2} \right)^{1/2} \sim 0.2.
		\label{eq:ndparam}
	\end{aligned}
\end{equation}
Here, $a$ is a measure of the strength of the gain fibre, $h$ is the loss due to absorption and the output coupler, $b$ is the ratio of the frequency shift due to the Kerr nonlinearity to the width of the pulse, and $s$ is the ratio of the effectiveness of the CFBG to the width of the pulse. Furthermore, notice that in this non-dimensional form, the modulation has no associated parameter, and each of the other processes in \eqref{eq:ndparam} have their own independent, non-dimensional parameter.

% Table of parameter values
 \begin{table*}[tbp]
	\centering
	\caption{Range of variation of various parameters.}
 	\label{tab:values}
 	\begin{tabular}{lcll}
 		\hline\noalign{\smallskip}
 		Parameter & Symbol & Value & Sources \\
 		\hline\noalign{\smallskip}
 		Absorption of Fibre & $\alpha$ & $10^{-4}$--$0.3\text{ m}^{-1}$  & \cite{burgoyneemail, shtyrina2017, tomlinson1984, usechak2005, yarutkina2013} \\
 		Fibre Dispersion & $\beta_2^f$ & $-50$--$50 \text{ ps}^2/ \text{km}$ & \cite{agrawal2002, agrawal2013, burgoyne2014, litchinitser1997, peng2018, yarutkina2013} \\
 		Fibre Nonlinearity & $\gamma$ & $0.001$--$0.01 \text{ W}^{-1} \text{m}^{-1}$ & \cite{agrawal2013, finot2008, usechak2005, yarutkina2013} \\
 		Grating Dispersion & $\beta_2^g L_D$ & $10$--$2000 \text{ ps}^2$ & \cite{agrawal2002, agrawal2013, burgoyne2014, li1998} \\
 		Length of Cavity & $L_T$ & $10$--$100 \text{ m}$ & \cite{burgoyneemail, peng2018, tamura1996} \\
 		Length of Fibre & $L_f$ & $0.15$--$1 \text{ m}$ & \cite{burgoyneemail} \\
 		Length of Gain Fibre & $L_g$ & $2$--$3 \text{ m}$ & \cite{burgoyne2014, peng2018, shtyrina2017, tamura1993,yarutkina2013} \\
 		Modulation Time & $T_M$ & $15$--$150 \text{ ps}$ & \cite{bohun2015, burgoyneemail, burgoyne2014} \\
 		Reflectivity of Optical Coupler & $R$ & $0.1$--$0.9$ & \cite{burgoyneemail, li1998, peng2018,  tamura1993, tamura1996, yamashita2009} \\
 		Saturation Energy & $\Es$ & $10^3$--$10^4 \text{ pJ}$ & \cite{burgoyneemail, usechak2005, yarutkina2013} \\
 		Small Signal Gain & $g_0$ & $1$--$10 \text{ m}^{-1}$ & \cite{burgoyneemail, yarutkina2013} \\
 		\noalign{\smallskip}\hline
 	\end{tabular}
 \end{table*}

@@

Now that we have the algebraic effect of each section of the cavity, we are ready to take the composition of the maps to give the effect of one round trip of the cavity. Thus, we must now consider the order of the components. As we are most interested in the output of the laser cavity, we shall start with the loss. We then pass the pulse through the CFBG followed by the modulator. The pulse then travels through the Er-doped fibre to be amplified, and finally we consider the effect of the nonlinearity. The nonlinearity must immediately follow the gain in this model---this is where the nonlinearity will have the largest effect since this is where the pulse is most energetic. Note, the permutation of the components is indeed important as the operators do not, in general, commute---unlike the models described in Section \ref{sec:discrete}. Moreover, we prefer the loss to follow the gain and nonlinearity as we wish to minimize the length over which the nonlinearity has an effect, and maximize the output power. Functionally we denote one round trip of the cavity by
\begin{align}
	\mathcal{L}(A) = F(G(M(D(L(A))))).
	\label{eq:order}
\end{align}
 The pulse after one complete circuit of the laser cavity is then returned back into the cavity to restart the process. A steady solution to this model is one in which the envelope and chirp are unchanged after traversing every component in the cavity---we are uninterested in the phase due to the nature of experimental measurements. That is, such that $\mathcal{L}(A) = A \textrm{e}^{i \phi}$---for some $\phi \in \mathbb{R}$. An example of the evolution of the envelope during one round trip of the laser cavity can be found in Fig. \ref{fig:cavityevo}.

\section{Results}
\label{sec:results}
We split the results into two subsections. In the following subsection we investigate the low nonlinearity limit, and in Section \ref{sec:nlresults} we look at the full nonlinear model.

\subsection{Linear Solution}
By neglecting the effect of the nonlinearity, that is, $b = 0$, a solution can be found analytically. We proceed by assuming the solution will take the form of a chirped Gaussian. There are a few reasons for this; the solution to the models presented in \cite{cutler1955, siegman1969, kuizenga1970a, martinez1984, martinez1985} were Gaussian, the equilibrium shape will be highly correlated to the modulation function, and since a Gaussian is a fixed point of the Fourier transform. Furthermore, this form is chosen because it resembles the envelope and linear chirp expected from the literature \cite{burgoyne2014, haus1975, haus1996, haus2000, usechak2005}.

>>>>

\section{Introduction}
\label{sec:intro}
Tuneable lasers have the ability to vary the frequency of their output by up to about 100 nanometres~\cite{bohun2015, burgoyne2010, yamashita2009}. This ability to vary the frequency leads to applications such as optical coherence tomography~\cite{bohun2015, burgoyne2014, yamashita2009}, coherent anti-Stokes Raman spectroscopy~\cite{burgoyne2014}, deep tissue multi-photon microscopy~\cite{chung2017, liu2017}, and diagnostics of ultrafast processes~\cite{burgoyne2014, silfvast2004}. Tuneable lasers are usually constructed in a ring with five main components: output coupler, chirped fibre Bragg grating (CFBG), modulator, Er-doped fibre, and pump laser. A typical ring laser cavity is depicted in Fig.~\ref{fig:cavity}. Due to the high power and short duration of the pulses, the Kerr nonlinearity plays a crucial role in the dynamics within the cavity.

Several effects arise within the cavity due to the interplay of dispersion, modulation, and the nonlinearity~\cite{bohun2015, coen1997, lapre2019, shao2019, woodward2018}. The two effects of most interest for this paper are wave breaking, and modulation instability. Wave breaking causes the leading edge of a pulse to be red shifted, and the trailing edge blue shifted (in the anomalous dispersion regime, the effect is reversed) to the point that a shock is developed~\cite{anderson1992, rothenberg1989a, rothenberg1989b, tomlinson1984, tomlinson1985}. The shifts in frequencies cause the pulse to become more rectangular in the frequency domain with a linear chirp over most of the pulse. In optics, wave breaking manifests from self-phase modulation (SPM) which is a direct effect of the Kerr nonlinearity causing the pulse to interfere with itself. New frequencies are then generated which induces the rectangular profile~\cite{agrawal2013, mollenauer1980, woodward2018}. Outside of optics, wave breaking occurs in other areas with nonlinear waves such as plasmas, heat propagation, transmission lines, and fluid dynamics~\cite{rothenberg1989b}. Modulation instability---which typically arises in the anomalous dispersion regime---is the other effect that will be of interest. However, in the presence of multiple frequencies, modulation instability can emerge in the normal dispersion regime as well through cross-phase modulation and four-wave mixing~\cite{agrawal1987, agrawal1989, agrawal2013, haelterman1992}. Modulation instability causes a wave to break into short pulses, and can cause a laser to lose coherence~\cite{agrawal1987, coen1997, haelterman1992}. Within a ring laser cavity wave breaking and modulation instability can become parasitic leading to an unstable and unsustainable pulse. This gives rise to the need of understanding the rich landscape of the parameter space, and determining design principles to guarantee the ring laser is stable and sustainable~\cite{bohun2015, burgoyneemail, finot2008, lapre2019, woodward2018}.

\begin{figure}[tbp]
	\centering
	\input{Cavity}
	\caption{Typical cavity of a fibre ring laser~\cite{burgoyne2014, chung2017, lapre2019, shao2019, tang2014}. The laser pulses travel clockwise around each loop.}
	\label{fig:cavity}
\end{figure}

In this paper, we will build an iterative map model for the evolution of a laser pulse as it travels though the cavity. In Section~\ref{sec:modelling}, we will briefly review previous mathematical models for ring lasers. Then, we will present our own iterative map model in Section~\ref{sec:model}, as well as non-dimensionalize the model and introduce four dimensionless parameters that govern the dynamics of the system. We will then examine the results of our model in Section~\ref{sec:results}---first solving the model analytically, by assuming the absence of the nonlinearity, then numerically solving the full nonlinear model. We will identify a sharp boundary in the parameter space separating the region of stability, and the region where modulation instability degrades the pulse. Finally, the concluding thoughts and possible ideas for future work are given in Section~\ref{sec:conclusion}.

\section{Modelling Efforts}
\label{sec:modelling}
The standard equation for studying nonlinear optics is the nonlinear Schr\"odinger equation (NLSE),
\begin{align}
	\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + i \gamma |A|^2 A.
	\label{eq:smallnlse}
\end{align}
Here, $A = A(T, z) : \mathbb{R}^2 \mapsto \mathbb{C}$ is the complex pulse amplitude, $\beta_2 \in \mathbb{R}$ is the second order dispersion, and $\gamma \in \mathbb{R}$ is the coefficient of nonlinearity. In practice, the NLSE, \eqref{eq:smallnlse}, lacks a few key terms, thus, it is often generalized by adding amplification and loss (occasionally higher order terms are added as well). These additions give the generalized nonlinear Schr\"{o}dinger equation (GNLSE)~\cite{agrawal2013, bohun2015, finot2008, peng2018, shtyrina2017, yarutkina2013},
	\begin{align}
	\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + i \gamma |A|^2 A + \frac{1}{2}g(A) A - \alpha A,
	\label{eq:nlse}
\end{align}
where $g(A)$ is an amplifying term due to the gain, and $\alpha \in \mathbb{R}^+$ is the loss due to scattering and absorption.

The GNLSE has many applications in nonlinear optics and fibre optic communications, however, in laser physics typically a modulation term is also added to ensure mode-locking, this yields the master equation of mode-locking~\cite{haus1984, haus1975, haus1986, haus1992, haus2000, tamura1996, usechak2005},
\begin{align}
	\pdiff{A}{z} &= - i \frac{\beta_2}{2}\pdiff[2]{A}{T} + i \gamma |A|^2 A + \frac{1}{2}g(A) A - \alpha A - M(T).
	\label{eq:meml}
\end{align}
No analytic solution is known for \eqref{eq:meml}---even after making the assumptions that the gain is constant, and the modulation is quadratic~\cite{haus1984, haus1975, haus1996}. However, by neglecting certain terms of \eqref{eq:meml} analytic solutions can be found~\cite{burgoyne2014, haus1975, haus1986, haus1991, haus1992, haus1996, tamura1996, usechak2005}. See Haus~\cite{haus2000} for a comprehensive description and history of the theory of mode-locked lasers.

\subsection{Discrete Models}
\label{sec:discrete}
Unfortunately, the master equation, \eqref{eq:meml}, is not completely representative of the underlying physics within our laser cavity. It is assumed in \eqref{eq:meml} that each process affects the pulse continuously within the cavity. As highlighted by Fig.~\ref{fig:cavity}, this can be a poor assumption. Within the cavity, each effect is localized to its corresponding component: almost all of the dispersion happens within the CFBG~\cite{agrawal2002}, the pulse is only amplified within the Erbium-doped fibre, \emph{et cetera}. Thus, a better model is one where we consider a simplified version of \eqref{eq:meml} for each component. Solving these simplified differential equations yields an algebraic expression for the effect of each individual component. The relations can then be functionally composed to give an iterative map for the effect of one round trip of the cavity.

Such a method was first proposed in 1955 by Cutler~\cite{cutler1955} while analyzing a microwave regenerative pulse generator. His method was adapted for mode-locked lasers in 1969 by Siegman and Kuizenga~\cite{kuizenga1970a, kuizenga1970b, kuizenga1970, siegman1969}. The effects of the nonlinearity would not be considered until Martinez \emph{et al.}~\cite{martinez1984, martinez1985} modelled passively mode-locked lasers. This issue has recently been readdressed by Burgoyne~\cite{burgoyne2014} in the literature for tuneable lasers. In each of these models the effect of each block is described by a transfer function.

Despite the development of discrete, block style models, several short-comings exist. None of the previous models have contained every component within a laser cavity---either the nonlinearity or the modulation have been omitted. Each component of a tuneable plays a pivotal role and the laser will not function correctly without the inclusion of all of the components. Another drawback is that the functional operations of some of the components used in their models are phenomenological. While these functions are chosen based on the observed output, they are not necessarily consistent with the underlying physics. Finally, the previous models have been unsuccessful in exhibiting the instabilities described in Section~\ref{sec:intro}.

\section{Model Derivation}
\label{sec:model}
Using the idea of a discrete model presented Section~\ref{sec:discrete}, we derive our model from the GNLSE, \eqref{eq:nlse}; however, in the case of the modulation we consider the exact functional form to be determined by the laser operator. To accomplish this we follow the method described in Section~\ref{sec:discrete}; we neglect all terms but one on the right side of \eqref{eq:nlse} and solve the simplified differential equations.

To proceed we must choose the form of the amplification term, $g(A)$. We assume the gain has the form
\begin{align}
	g(A) &= \frac{g_0}{1 + E / \Es},& E &= \int_{-\infty}^\infty |A|^2 \, \df T,
	\label{eq:energy}
\end{align}
where $g_0$ is a small signal gain, $E$ is the energy of the pulse, and $\Es$ is the energy at which the gain begins to saturate~\cite{haus1984, shtyrina2017, silfvast2004}. This reduces \eqref{eq:nlse} to
\begin{align}
	\pdiff{A}{z} &= \frac{g_0 A}{2 \left( 1 + E / \Es \right)}.
\end{align}
By assuming the incoming energy is $E$ and the energy after travelling through the gain fibre of length $L_g$ is $E_{\text{out}}$, we find the energy after amplification is
\begin{align}
	\label{eq:engain}
	E_{\text{out}} &= \Es W \left( \frac{E}{\Es} \exp( E / \Es ) \exp( g_0 L_g ) \right),
\end{align}
where $W$ is the Lambert $W$ function. Rewriting \eqref{eq:engain} in terms of the pulse gives
\begin{align}
	G(A) &= \left( \frac{\Es}{E} W \left( \frac{E}{\Es} \exp( E / \Es ) \exp( g_0 L_g ) \right) \right)^{1/2} A
	\label{eq:gainform}
\end{align}
for the amplification of the pulse within the Er-doped fibre. Repeating this process for the loss, dispersion, and nonlinearity yields
\begin{align}
	L(A) &= (1 - R) \exp( - \alpha L_T )A, \\
	D(A) &= \FTi{\exp( i \omega^2 L_D\beta_2/2 ) \FT{A}}, \\
	F(A) &= A \exp( i \gamma |A|^2 L_f ),
\end{align}
where $R$ is the reflectivity of the output coupler (depending on the layout of the laser cavity the loss may instead take the form $L(A) = R \exp( - \alpha L_T )A$), $L_T$ is the total length of the laser circuit, $L_D$ is the length of the dispersive medium, $L_f$ is the length of fibre between the amplifier and the output coupler, and $\mathcal{F}$ denotes the Fourier transform. Finally, we must consider the modulation. The modulation is considered to be applied externally in which ever way the operator sees fit. For simplicity the representation is taken as the Gaussian
\begin{align}
	M(A) &= \exp( -T^2 / 2 T_M^2 ) A,
	\label{eq:modform}
\end{align}
where $T_M$ is the characteristic width of the modulation. This assumption is not particularly restrictive. Calcaterra and Boldt have shown that Gaussians can form a basis for $L^2(\mathbb{R})$~\cite{calcaterra2008a}. Therefore, we can approximate any modulation function in $L^2(\mathbb{R})$ by a sum of \eqref{eq:modform}.

\subsection{Non-Dimensionalization}
Non-dimensionalization is a technique used in mathematical modelling to better understand the relative importance of different processes within a system. In essence, we scale all variables by typical values of the system based on the geometry and boundary conditions (see, for example,~\cite{howison2005}). Doing so `factors' the dimensional units out of the problem, leaving only dimensionless equations with dimensionless parameters. The magnitude of the dimensionless parameters characterizes the dominant dynamics of the system---much like how fluid flow can be described with only the Reynolds number. The structure of each process in the laser can be better understood by re-scaling the time, energy, and amplitude by the convenient scalings:
\begin{align}
	T &= T_M \widetilde{T},& E &= \Es \widetilde{E},& A &= \left( \frac{\Es}{T_M} \right)^{1/2} \widetilde{A}.
\end{align}

Revisiting the process maps \eqref{eq:gainform}--\eqref{eq:modform} yields the new, non-dimensional, mappings---after dropping the tildes---
\begin{align}
	G(A) &= \left(E^{-1} W \left( a E \textrm{e}^{E}\right) \right)^{1/2} A, & F(A) &= A \exp( i b |A|^2), \nonumber \\
	D(A) &= \FTi{\exp( i s^2 \omega^2 ) \FT{A}}, & L(A) &= h A, \label{eq:effects} \\
	M(A) &= \exp( -T^2 / 2 ) A, \nonumber
\end{align}
with the four dimensionless parameters, as defined by the values in Table~\ref{tab:values},
\begin{equation}
	\begin{aligned}
		a &= \exp( g_0 L_g ) \sim 8 \times 10^3, & 
		h &= (1 - R) \exp( - \alpha L_T) \sim 0.04, \\
		b &= \gamma L_f \frac{\Es}{T_M} \sim 1, & s &= \left( \frac{\beta_2 L_D}{2 T_M^2} \right)^{1/2} \sim 0.2.
		\label{eq:ndparam}
	\end{aligned}
\end{equation}
Here, $a$ is a measure of the strength of the gain fibre, $h$ is the loss due to absorption and the output coupler, $b$ is the ratio of the frequency shift due to the Kerr nonlinearity to the width of the pulse, and $s$ is the ratio of the effectiveness of the CFBG to the width of the pulse. Furthermore, notice that in this non-dimensional form, the modulation has no associated parameter, and each of the other processes in \eqref{eq:ndparam} have their own independent, non-dimensional parameter.

% Table of parameter values
 \begin{table*}[tbp]
	\centering
	\caption{Range of variation of various parameters.}
 	\label{tab:values}
 	\begin{tabular}{lcll}
 		\hline\noalign{\smallskip}
 		Parameter & Symbol & Value & Sources \\
 		\hline\noalign{\smallskip}
 		Absorption of Fibre & $\alpha$ & $10^{-4}$--$0.3\text{ m}^{-1}$  &~\cite{burgoyneemail, shtyrina2017, tomlinson1984, usechak2005, yarutkina2013} \\
 		Fibre Dispersion & $\beta_2^f$ & $-50$--$50 \text{ ps}^2/ \text{km}$ &~\cite{agrawal2002, agrawal2013, burgoyne2014, litchinitser1997, peng2018, yarutkina2013} \\
 		Fibre Nonlinearity & $\gamma$ & $0.001$--$0.01 \text{ W}^{-1} \text{m}^{-1}$ &~\cite{agrawal2013, finot2008, usechak2005, yarutkina2013} \\
 		Grating Dispersion & $\beta_2^g L_D$ & $10$--$2000 \text{ ps}^2$ &~\cite{agrawal2002, agrawal2013, burgoyne2014, li1998} \\
 		Length of Cavity & $L_T$ & $10$--$100 \text{ m}$ &~\cite{burgoyneemail, peng2018, tamura1996} \\
 		Length of Fibre & $L_f$ & $0.15$--$1 \text{ m}$ &~\cite{burgoyneemail} \\
 		Length of Gain Fibre & $L_g$ & $2$--$3 \text{ m}$ &~\cite{burgoyne2014, peng2018, shtyrina2017, tamura1993,yarutkina2013} \\
 		Modulation Time & $T_M$ & $15$--$150 \text{ ps}$ &~\cite{bohun2015, burgoyneemail, burgoyne2014} \\
 		Reflectivity of Optical Coupler & $R$ & $0.1$--$0.9$ &~\cite{burgoyneemail, li1998, peng2018,  tamura1993, tamura1996, yamashita2009} \\
 		Saturation Energy & $\Es$ & $10^3$--$10^4 \text{ pJ}$ &~\cite{burgoyneemail, usechak2005, yarutkina2013} \\
 		Small Signal Gain & $g_0$ & $1$--$10 \text{ m}^{-1}$ &~\cite{burgoyneemail, yarutkina2013} \\
 		\noalign{\smallskip}\hline
 	\end{tabular}
 \end{table*}

@@

Now that we have the algebraic effect of each section of the cavity, we are ready to take the composition of the maps to give the effect of one round trip of the cavity. Thus, we must now consider the order of the components. As we are most interested in the output of the laser cavity, we shall start with the loss. We then pass the pulse through the CFBG followed by the modulator. The pulse then travels through the Er-doped fibre to be amplified, and finally we consider the effect of the nonlinearity. The nonlinearity must immediately follow the gain in this model---this is where the nonlinearity will have the largest effect since this is where the pulse is most energetic. Note, the permutation of the components is indeed important as the operators do not, in general, commute---unlike the models described in Section~\ref{sec:discrete}. Moreover, we prefer the loss to follow the gain and nonlinearity as we wish to minimize the length over which the nonlinearity has an effect, and maximize the output power. Functionally we denote one round trip of the cavity by
\begin{align}
	\mathcal{L}(A) = F(G(M(D(L(A))))).
	\label{eq:order}
\end{align}
 The pulse after one complete circuit of the laser cavity is then returned back into the cavity to restart the process. A steady solution to this model is one in which the envelope and chirp are unchanged after traversing every component in the cavity---we are uninterested in the phase due to the nature of experimental measurements. That is, such that $\mathcal{L}(A) = A \textrm{e}^{i \phi}$---for some $\phi \in \mathbb{R}$. An example of the evolution of the envelope during one round trip of the laser cavity can be found in Fig.~\ref{fig:cavityevo}.

\section{Results}
\label{sec:results}
We split the results into two subsections. In the following subsection we investigate the low nonlinearity limit, and in Section~\ref{sec:nlresults} we look at the full nonlinear model.

\subsection{Linear Solution}
By neglecting the effect of the nonlinearity, that is, $b = 0$, a solution can be found analytically. We proceed by assuming the solution will take the form of a chirped Gaussian. There are a few reasons for this; the solution to the models presented in~\cite{cutler1955, siegman1969, kuizenga1970a, martinez1984, martinez1985} were Gaussian, the equilibrium shape will be highly correlated to the modulation function, and since a Gaussian is a fixed point of the Fourier transform. Furthermore, this form is chosen because it resembles the envelope and linear chirp expected from the literature~\cite{burgoyne2014, haus1975, haus1996, haus2000, usechak2005}.
